---
title: "Week 7 - HTML and JSON"
author: "Kristoff Oliphant"
format: html
editor: visual
---

## Introduction

It is now week 7 and the assignment this week will involve HTML and JSON. This assignment is considered a warm up exercise to help us get familiar with the HTML and JSON file formats, and using packages to read these data formats for downstream use in R data frames

## Planned Workflow

For my planned workflow, I selected three books involving data science. I chose three books with multiple authors and for each book, I recorded the title, authors, and two to three additional attributes such as the publication year, publisher and ISBN. With the information from the books I manually created two textedit files, one HTML file containing a table with the book data and changed its file type to html. I also used a textedit file and manually created a JSON file with the same data as the html but with the file type .json. I’ll open and load both the html and json files in separate R dataframes, and use tidyverse and dplyr to manipulate the data into a readable format in R. I’ll load the data using rvest to read the html and jsonlite for the json file, ensuring it’s formatted consistently for an identical comparison.

## Anticipated Challenges

A challenge I anticipate facing is making sure the syntax in both files are correct to be loaded correctly into R. I manually created these and if I didn’t follow the logic correctly in each of its respective files, my code in R has a high chance of failing if it's incorrect. Since both files use a unique syntax there’s a fair chance both can load into R in different ways that don’t match each other.

```{r}
library(rvest)
library(jsonlite)
library(tidyverse)
```

```{r}
html_raw <- read_html("https://raw.githubusercontent.com/Kristoffgit/data607_week7/refs/heads/main/books.html")
html_df <- html_raw %>%
  html_table() %>%
  .[[1]]
```

```{r}
print(html_df)
```

```{r}
json_raw <- fromJSON("https://raw.githubusercontent.com/Kristoffgit/data607_week7/refs/heads/main/books.json")
json_df <- as.data.frame(json_raw$books)
```

```{r}
print(json_df)
```

```{r}
str(html_df)
str(json_df)
```

```{r}
json_df <- json_df %>%
  mutate(Authors = sapply(Authors, paste, collapse = ", ")) %>%
  rename(Author = Authors)
```

```{r}
html_df <- html_df %>% arrange(Title)
json_df <- json_df %>% arrange(Title)
```

```{r}
is_identical <- identical(html_df, json_df)
```

```{r}
print(is_identical)
```

```{r}
if(!is_identical) {
  all.equal(html_df, json_df)
}
```

**Conclusion**

After comparing both dataframes, the identical function returned that it was FALSE. From what I notice, the data frames have some differences in their object attributes and string formatting. The string mismatch is because Cetinkaya-Rundel's name has a special character in C for JSON and has a regular C in HTML.
